---
title: "final-project"
output: pdf_document
---

```{r setup, include=FALSE, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(knitr)
```

# Proliferation of Regime Types Across Time

Before I get into this, I just want to communicate that I do not know the actual parameters of this assignment, both for now and for the final project as a whole. I actually thought this was due today, not one week ago.

The purpose of this project is to observe the changes in regime type prevalence relative to the strength of certain regimes, using a temporal and geographic lens in order to better understand the effect that major regime changes can have on surrounding regions. In order to understand this, the study is going to look at a number of cases over the course of the 20th century, with the Soviet Union and single party regimes being an example. The analysis will be conducted over the course of multiple year ranges and geographic locations; it will consist of a number of cases and analyses. 

The data comes from a study published on March 14th, 2018 by Carsten Anckar and Cecilia Fredriksson in the Journal of European Political Science. It classifies political regimes from 1800 to 2016, using classifications which include a country's democratic/non-democratic status, a country's regime type (presidentialist, parliamentarian, personalist rule, etc.) and whether a country's head of state is popularly elected. The data in question is updated over the course of years, meaning that countries' regimes change over time during the dataset; consequently, the observed slopes are likely to be quite gentle. 

\newpage

```{r data, echo = FALSE}

data <- readxl::read_xlsx("data/41304_2018_149_MOESM2_ESM (1).xlsx")

# This was to make viewing the countries on the list easier. 

y2000 <- data %>%
  filter(year == "2000")
view(y2000)

# This makes regression across the government types easier. 

data <- data %>%
  mutate(parliamentarism = ifelse(regimenarrowcat == 0, 1, 0),
         semipresidentialism = ifelse(regimenarrowcat == 1, 1, 0),
         presidentialism = ifelse(regimenarrowcat == 2, 1, 0),
         semimonarchy = ifelse(regimenarrowcat == 3, 1, 0),
         singleparty = ifelse(regimenarrowcat == 4, 1, 0),
         multipartyauthoritarian = ifelse(regimenarrowcat == 5, 1, 0),
         personalist = ifelse(regimenarrowcat == 6, 1, 0),
         military = ifelse(regimenarrowcat == 7, 1, 0),
         absolutemonarchy = ifelse(regimenarrowcat == 8, 1, 0),
         monarchicoligarchy = ifelse(regimenarrowcat == 9, 1, 0),
         otheroligarchy = ifelse(regimenarrowcat == 10, 1, 0),
         missing = ifelse(regimenarrowcat == 99, 1, 0))

# This allows for an easier geographic perspective. 

data <- data %>%
  mutate(asian = ifelse(
    ccode >= 630, 1, 0),
    african = ifelse(
      ccode >= 402 & ccode < 630, 1, 0),
    european = ifelse(
      ccode >= 200 & ccode <= 395, 1, 0),
    american = ifelse(
      ccode >= 2 & ccode < 200, 1, 0))



```

```{r ussr, echo = FALSE}

#data_ussr

data_ussr <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 17) %>%
  select(year_new, singleparty)

lm_ussr <- lm(singleparty ~ year_new, data = data_ussr)

summary(lm_ussr)$adj.r.squared

#data_ussr_alt

data_ussr_alt <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 50) %>%
  select(year_new, singleparty)

lm_ussr_alt <- lm(singleparty ~ year_new, data = data_ussr_alt)

lm_ussr_alt_sq <- lm(singleparty ~ year_new +
                    I(year_new^2), data = data_ussr_alt)

summary(lm_ussr_alt)$adj.r.squared

summary(lm_ussr2_sq)$adj.r.squared

lm_ussr_alt_sq_values <- data_ussr_alt %>% select(year_new, singleparty)
lm_ussr_alt_sq_values$ussr_alt_sq_values <-
  predict(lm_ussr_alt_sq, newdata = lm_ussr_alt_sq_values)

# The alt was designed to better compare with ussr2

#data_ussr2

data_ussr2 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 17, european == 1) %>%
  select(year_new, singleparty)

lm_ussr2 <- lm(singleparty ~ year_new, data = data_ussr2)

lm_ussr2_sq <- lm(singleparty ~ year_new +
                    I(year_new^2), data = data_ussr2)

summary(lm_ussr2)$adj.r.squared

summary(lm_ussr2_sq)$adj.r.squared

lm_ussr2_sq_values <- data_ussr2 %>% select(year_new, singleparty)
lm_ussr2_sq_values$ussr2_sq_values <-
  predict(lm_ussr2_sq, newdata = lm_ussr2_sq_values)

# First regression includes all countries, second only includes Europe. 

#data_ussr3

data_ussr3 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 50, european == 1) %>%
  select(year_new, singleparty)

lm_ussr3 <- lm(singleparty ~ year_new, data = data_ussr3)

lm_ussr3_sq <- lm(singleparty ~ year_new +
                    I(year_new^2), data = data_ussr3)

summary(lm_ussr3)$adj.r.squared

summary(lm_ussr3_sq)$adj.r.squared

lm_ussr3_sq_values <- data_ussr3 %>% select(year_new, singleparty)
lm_ussr3_sq_values$ussr3_sq_values <-
  predict(lm_ussr3_sq, newdata = lm_ussr3_sq_values)

#data_ussr3_alt

data_ussr3_alt <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 50, european == 1 | asian == 1) %>%
  select(year_new, singleparty)

lm_ussr3_alt <- lm(singleparty ~ year_new, data = data_ussr3_alt)

lm_ussr3_alt_sq <- lm(singleparty ~ year_new +
                    I(year_new^2), data = data_ussr3_alt)

summary(lm_ussr3_alt)$adj.r.squared

summary(lm_ussr3_alt_sq)$adj.r.squared

lm_ussr3_alt_sq_values <- data_ussr3_alt %>% select(year_new, singleparty)
lm_ussr3_alt_sq_values$ussr3_alt_sq_values <-
  predict(lm_ussr3_alt_sq, newdata = lm_ussr3_alt_sq_values)

# Third regression changes the baseline year. 

#data_ussr4

data_ussr4 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 75, european == 1) %>%
  select(year_new, singleparty)

lm_ussr4 <- lm(singleparty ~ year_new, data = data_ussr4)

summary(lm_ussr4)$adj.r.squared

#data_ussr5

data_ussr5 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new >= 87, european == 1) %>%
  select(year_new, singleparty)

lm_ussr5 <- lm(singleparty ~ year_new, data = data_ussr5)

summary(lm_ussr5)$adj.r.squared

#data_ussr5_alt

data_ussr5_alt <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new >= 87, european == 1 | asian == 1) %>%
  select(year_new, singleparty)

lm_ussr5_alt <- lm(singleparty ~ year_new, data = data_ussr5_alt)

summary(lm_ussr5_alt)$adj.r.squared

# The fourth and fifth regressions also change baseline year.

# Interestingly, the correlation shifts significantly over time. 

# Regime change 1917 - 1950

data_ussr2_test <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 17, european == 1) %>%
  select(year_new, singleparty) %>%
  mutate(regime_type1917 = singleparty)

data_ussr3_test <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 50, european == 1) %>%
  select(year_new, singleparty) %>%
  mutate(regime_type1950 = singleparty)

data_ussr_regime_change1950 <- full_join(data_ussr2_test, data_ussr3_test, by = "year_new") %>%
  na.omit() %>%
  mutate(regime_change = ifelse(regime_type1917 != regime_type1950,
                                1, 0))
lm_ussr_regime_change1950 <- lm(regime_change ~ year_new,
  data = data_ussr_regime_change1950)

summary(lm_ussr_regime_change1950)$adj.r.squared

# Regime change 1950 - 1987

data_ussr4_test <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 87, european == 1) %>%
  select(year_new, singleparty) %>%
  mutate(regime_type1987 = singleparty)

data_ussr_regime_change1987 <- full_join(data_ussr3_test, data_ussr4_test, by = "year_new") %>%
  na.omit() %>%
  mutate(regime_change = ifelse(regime_type1950 != regime_type1987,
                                1, 0))
lm_ussr_regime_change1987 <- lm(regime_change ~ year_new,
  data = data_ussr_regime_change1987)

summary(lm_ussr_regime_change1987)$adj.r.squared

# New Idea for Regime Change Graph

data_ussr2_test_2 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new == 17, european == 1) %>%
  select(year_new, singleparty, country) %>%
  mutate(regime_type1917 = singleparty)

data_ussr3_test_2 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new == 50, european == 1) %>%
  select(year_new, singleparty, country) %>%
  mutate(regime_type1950 = singleparty)

data_ussr_regime_change1950_2 <- 
  full_join(data_ussr2_test_2, data_ussr3_test_2, by = "country") %>%
  na.omit() %>%
  mutate(regime_change = ifelse(regime_type1917 != regime_type1950,
                                1, 0)) %>%
  select(country, regime_change)

data_ussr_regime_change1950_2.5 <- 
  full_join(data_ussr2_test_2, data_ussr3_test_2, by = "country") %>%
  na.omit() %>%
  mutate(regime_change = ifelse(regime_type1917 != regime_type1950,
                                1, 0))

# Same Idea but Different Years

data_ussr4_test_2 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new == 91, european == 1) %>%
  select(year_new, singleparty, country) %>%
  mutate(regime_type1987 = singleparty)

data_ussr_regime_change1950_3 <- 
  full_join(data_ussr3_test_2, data_ussr4_test_2, by = "country") %>%
  na.omit() %>%
  mutate(regime_change = ifelse(regime_type1950 != regime_type1987,
                                1, 0)) %>%
  select(country, regime_change)

```

## The Case of the USSR

(there will be more graphs in the final version)

The below two graphs show two ways of interpreting the potential spread of single party regimes following the emergence of the Soviet Union in 1917. The independent variable is the year, while the dependent variable is the prevalence of single party regimes. The difference between the two graphs is what region they cover: the first graph includes all the world's countries, whereas the second graph limits the analysis to only Europe. All of the lines in these graphs were modified by a 30x multiplier to better reflect the data; it should be notified that this multiplier only exists to make the lines more visible and, though the numbers are proportional, the axes should not be considered when observing the lines. 

The first graph depicts the distribution of single party regimes in the world over the course of the 20th century. The data begins in 1917, the year of the Russian Revolution. It is important to note that the prevalence of single party regimes does not change significantly following 1917, but rather following 1945. 

The second graph depicts the distribution of single party regimes specifically in Europe over the 20th century. In this graph, there are two lines: a linear line and a quadratic line. The linear line's r-squared is 0.006696221, whereas the r-squared the quadratic line is 0.08567992, meaning the quadratic line better fits the data. The quadratic model reflects the rise and fall of the single party regime type over the course of the 20th century, as the USSR gained prevalence and fell over the course of 80 years.

A simple view at these graphs and at their correlations will communicate that this comparison is not strong. The linear models for both graphs do not reflect the drastic changes in the population of single party regimes around the years 1950 and 1991, and even the quadratic model featured in the European graph only slightly reflects this. Consequently, while the years since 1917 has some predictive power on the prevalence of single party regimes both in the world and particularly in Europe, our analysis can be improved.

```{r ussrgeom, echo = FALSE}

ussr_geom <- ggplot(data = data_ussr, mapping = 
                      aes(x = year_new, y = singleparty)) +
  geom_col(aes(y = singleparty)) +
  geom_abline(intercept = .1431769 * 30, slope = -0.0006335 * 30) +
  ylim(0, 40) + 
  xlim(15, 100) + 
  theme_bw() +
  labs(title = "Single Party Regimes in the 20th century",
       x = "Years since 1900",
       y = "Single Party Regimes")

# Both of these lines have been modified by a flat 30x multiplier
# in order to better compare the graphs with the data set
# this multiplier is arbitrary

ussr2_geom <- ggplot(data = data_ussr2, mapping = 
                      aes(x = year_new, y = singleparty)) +
  geom_col(aes(y = singleparty)) +
  geom_line(aes(y = 30*lm_ussr2_sq_values$ussr2_sq_values), color = "red") +
  geom_abline(intercept = .1859563 * 30, slope = - 0.0009342* 30,
              color = "blue") +
  ylim(0, 10) + 
  xlim(15, 100) + 
  theme_bw() +
  labs(title = "Single Party Regimes in 20th century Europe",
       x = "Years since 1900",
       y = "Single Party Regimes in Europe")

ussr_geom
ussr2_geom



```

\newpage

### Changing the Year: What Happens?

  Given that this analysis is attempting to use the year since a certain event as a predictor of a certain regime type in order to determine to what extent a demonstration effect exists for regime change, changing the year at which the analysis starts, i.e. what year the event occurs, would be a prudent decision. The below graphs do that, with certain graphs also expanding the geographic region beyond Europe to see if there is a greater correlation in Europe and Asia (the two continents which contained the USSR) than in Europe alone. 
  
  As a result of the method by which the lines of fit were calculated, additional changes were required in order to make them better reflect the column graph in the background. The first graph makes use of a 100x multiplier, while the subsequent graphs use a 30x multiplier; these multipliers were not set based on a rigorous statistical method and, as such, it should be understood that the lines do not correspond precisely to the column graphs behind them. To reiterate, the lines are there to communicate the trend in the data, but are not precise to the column graph itself. 
  
    The first graph shows the concentration of single party regimes worldwide from the year 1950 to the year 2000. The graph shows a steady increase from 1950 to 1960, stagnation through the mid-sixties to the mid-seventies, and then a jump following approximately 1975, peaking at around 1978. Following that, the concentration of single party regimes seems to decline, with it dropping dramatically in the early 1990s. The line of best fit for this set of data shows a steady decline beginning at a moderate height; the quadratic curve for this model begins at a lower amount, slightly rises to peak around 1962, and then gently curves down toward 2000, finishing slightly above the linear fit. The r-squared value for the linear fit is 0.03463, while the correlation for the quadratic fit is 0.04009, meaning that the quadratic fit is not particularly better at predicting this relationship than the linear fit. 
    
    The second graph features the data from Europe. It begins high at about 8 single party regimes in 1950, goes up to about 9 around 1953, and stays at that level until the late 1980s, in which it begins to fall dramatically. The fits for this set of data are much steeper than in the previous graph; the linear and quadratic fits begin relatively higher when compared to the previous graph, but also fall much faster; this reflects the swift pace at which single party regimes disappeared in Eurppe as the Soviet Union weakened. The r-squared for the linear fit is 0.1066, while the r-squared for the quadratic fit is 0.1094; these fits essentially hold the same predictive power over this period. These r-squared are notably higher than those worldwide, suggesting that the geographic connection may play an important role in the effectiveness of the demonstration effect. 
    
    The third graph includes the same years as the previous two graphs, but focuses on Europe and Asia, as opposed to the entire world or just Europe. The graph mixes the patterns of the previous two graphs, beginning at a moderate height, increasing somewhat in the mid-seventies, and falling off significantly in the late 1980s. The linear fit begins much higher than the quadratic fit and follows a much steeper trend downward; the two fits intersect in the late 1980s, immediately before the drop. The r-squared value of the linear fit is 0.04610 while the r-squared value of the quadratic fit is 0.04824; as such, neither fit is particularly stronger than other. 
    
    When comparing the first three graphs, it becomes clear that the model with the strongest predictive power is that featured in the second graph, which pertains but to Europe. As such, our analysis of single party regime types in the 20th century suggests a geographic component to the demonstration effect, wherein a country's proximity to the epicenter of a certain regime type, in this case single party regimes emerging out of the USSR, increases its likelihood toward adopting such a regime type. However, it is also clear that, based on the low r-squared values, more goes into whether a country selects a given regime type than the demonstration effect. At its most robust thus far, the demonstration effect, measured in years since 1950, has accounted for approximately 11% of the variation in single party regimes in Europe. Though this number is respectable, its smallness suggests that a number of other factors exist which play a role in changing regime type. 
    
    Moving onto the last three graphs which cover the years since 1975 in Europe, the years since 1987 in Europe, and the years in 1987 in Europe and Asia, there is some sign of stronger measure. The fourth graph follows a similar pattern to the second graph; however, its r-squared value is slightly higher than the second graph at 0.1286. Interestingly, this graph's steeper, downward trend suggests an potential unraveling of the demonstration effect, where the Soviet Union's failure to maintain its coercive control over the various European states result in the collapse of the single party system in Europe as a whole. 
    
  The fifth and sixth graphs are much less interesting than the previous four. The fifth graph consists of only three years, those after 1987, and shows the drop that occurred in the late 1980s from a much closer perspective. Due to the limited sample size afforded to the last graph, the line of best fit does not capture the pattern here particularly well, with an r-squared value of just 0.0615. The sixth graph does not fare much better, with an r-squared of 0.0156. The graph starts slightly higher than the fifth one and retains approximately 3 single party regimes following 1991, which would account for those regimes in Asia; the decreased predictive power of the model once again suggests the role of geography in the demonstration effect. 
```{r ussr2, echo = FALSE}

ussralt_geom <- ggplot(data = data_ussr_alt, mapping = 
                      aes(x = year_new, y = singleparty)) +
  geom_col(aes(y = singleparty)) +
  geom_line(aes(y = 100*lm_ussr_alt_sq_values$ussr_alt_sq_values),
            color = "red") +
  geom_abline(intercept = .37460 * 100, slope = -0.00312 * 100) +
  ylim(0, 40) + 
  xlim(50, 100) + 
  theme_bw() +
  labs(title = "Single Party Regimes in the mid-20th century",
       x = "Years since 1900",
       y = "Single Party Regimes")

ussr3_geom <- ggplot(data = data_ussr3, mapping = 
                      aes(x = year_new, y = singleparty)) +
  geom_col(aes(y = singleparty)) +
  geom_line(aes(y = 30*lm_ussr3_sq_values$ussr3_sq_values), color = "red") +
  geom_abline(intercept = .642678 * 30, slope = -0.005878 * 30) +
  ylim(0, 10) + 
  xlim(50, 100) + 
  theme_bw() +
  labs(title = "Single Party Regimes in mid-20th century Europe",
       x = "Years since 1900",
       y = "Single Party Regimes in Europe")

ussr3alt_geom <- ggplot(data = data_ussr3_alt, mapping = 
                      aes(x = year_new, y = singleparty)) +
  geom_col(aes(y = singleparty)) +
  geom_line(aes(y = 30*lm_ussr3_alt_sq_values$ussr3_alt_sq_values), color = "red") +
  geom_abline(intercept = .642678 * 30, slope = -0.005878 * 30) +
  ylim(0, 20) + 
  xlim(50, 100) + 
  theme_bw() +
  labs(title = "Single Party Regimes in mid-20th century Europe and Asia",
       x = "Years since 1900",
       y = "Single Party Regimes in Europe and Asia")

ussr4_geom <- ggplot(data = data_ussr4, mapping = 
                      aes(x = year_new, y = singleparty)) +
  geom_col(aes(y = singleparty)) +
  geom_abline(intercept = .870932 * 30, slope = -0.008172 * 30) +
  ylim(0, 10) + 
  xlim(75, 100) + 
  theme_bw() +
  labs(title = "Single Party Regimes in Mid-Late 20th Century Europe",
       x = "Years since 1900",
       y = "Single Party Regimes in Europe")

ussr5_geom <- ggplot(data = data_ussr5, mapping = 
                      aes(x = year_new, y = singleparty)) +
  geom_col(aes(y = singleparty)) +
  geom_abline(intercept = .365125 * 30, slope = -0.003399 * 30) +
  ylim(0, 10) + 
  xlim(87, 91) + 
  theme_bw() +
  labs(title = "Single Party Regimes in Late 20th century Europe",
       x = "Years since 1900",
       y = "Single Party Regimes in Europe")

ussr5alt_geom <- ggplot(data = data_ussr5_alt, mapping = 
                      aes(x = year_new, y = singleparty)) +
  geom_col(aes(y = singleparty)) +
  geom_abline(intercept = .291841 * 30, slope = -0.002451 * 30) +
  ylim(0, 20) + 
  xlim(87, 100) + 
  theme_bw() +
  labs(title = "Single Party Regimes in Late 20th Century Europe and Asia",
       x = "Years since 1900",
       y = "Single Party Regimes in Europe and Asia")

ussralt_geom
ussr3_geom
ussr3alt_geom
ussr4_geom
ussr5_geom
ussr5alt_geom

```

\newpage

  Another way to measure the demonstration effect is to look at regime changes that have occurred within a given time frame, and examine how that may be related to a certain event. Unfortunately, the data available to this analysis makes performing this in the form of a linear regression over impracticable but, given that this information could still potentially be valuable, it shall be included as an addendum to the USSR section in the form of two tables. The two tables show a given European country in the left column and whether it changed from and to a single party regime between the set years. It should be noted that these tables measure whether the regime changed to or from being a single party regime in 1950 relative to 1917, and thus does not account for variation in between those years. This means that, if a single party regime were to emerge in 1924 and collapse in 1945, that  would not be accounted for by these tables; these tables also do not include certain European countries, as there was not data for them in the years covered. 

```{r ussr_regime_change, echo = FALSE}

# ussr_regime_change1950_geom <- ggplot(data = data_ussr_regime_change1950,
     #                                 mapping = aes(x = year_new,
    #                                            y = regime_change)) +
  # geom_col(aes(y = regime_change)) +
  # geom_abline(intercept = .043916 * 30, slope = 0.008803 * 30) +
  # ylim(0, 100) + 
  # xlim(50, 75) + 
  # theme_bw() +
  # labs(title = "Single Party Regimes in Late 20th Century Europe and Asia",
       # x = "Years since 1900",
       # y = "Regime Change in Europe*",
       # subtitle = "*Regime change number includes years after regime change")

#ussr_regime_change1987_geom <- ggplot(data = data_ussr_regime_change1987,
     #                                 mapping = aes(x = year_new,
     #                                           y = regime_change)) +
  # geom_col(aes(y = regime_change)) +
 # geom_abline(intercept = .043916 * 30, slope = 0.008803 * 30) +
 # ylim(0, 20) + 
  # xlim(87, 100) + 
  # theme_bw() +
  # labs(title = "Single Party Regimes in Late 20th Century Europe and Asia",
    #   x = "Years since 1900",
     #  y = "Regime Change in Europe*",
      # subtitle = "*Regime change number includes years after regime change")

knitr::kable(data_ussr_regime_change1950_2, col.names = 
               c("Country", "Regime Change between 1917 and 1950"
               ))

knitr::kable(data_ussr_regime_change1950_3, col.names = 
               c("Country", "Regime Change between 1950 and 1991"
               ))


```

## The Case of China

```{r china, echo = FALSE}

data_china <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 49) %>%
  select(year_new, singleparty)

lm_china <- lm(singleparty ~ year_new, data = data_china)

summary(lm_china)$adj.r.squared

data_china2 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 49, asian == 1) %>%
  select(year_new, singleparty)

lm_china2 <- lm(singleparty ~ year_new, data = data_china2)

summary(lm_china2)$adj.r.squared

data_china3 <- data %>%
  mutate(year_new = year - 1900) %>%
  filter(year_new > 75, asian == 1) %>%
  select(year_new, singleparty)

lm_china3 <- lm(singleparty ~ year_new, data = data_china3)

summary(lm_china3)$adj.r.squared


``` 

## Code

```{r print-code, ref.label=knitr::all_labels(), echo = TRUE, eval=FALSE}
## This command will print all of your named chunks
```



